Why aren't significance tests commonly used for linear regression diagnostics?

### 2022-09-21 paper_comparison/paper_comparison/paper_comparison.Rmd
update exp design for exp I

---
title: |
  Why aren't significance tests commonly used for linear regression diagnostics?
type: ARTICLE TEMPLATE
author:
  - name: Weihao Li
    affil: a
    email: weihao.li@monash.edu
  - name: Dianne Cook
    affil: a
    email: dicook@monash.edu
  - name: Emi Tanaka
    affil: a
    email: emi.tanaka@monash.edu
affiliation:
  - num: a
    address: |
      Department of Econometrics and Business Statistics, Monash University, Clayton, VIC, Australia
bibliography: paper.bib
abstract: |
  Abstract to fill.
keywords: |
  data visualization; visual inference; hypothesis testing; residual plots;
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
  \newcommand*{\Perm}[2]{{}^{#1}\!P_{#2}}%
output: rticles::tf_article
---

@@

During the experiment, every subject was presented with a block of 20 lineups. And for every lineup, the subject was asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots were from others. If there was no noticeable difference between plots in a lineup, subjects were permitted to select zero plots without providing the reason. No subject was shown the same lineup twice. Information about preferred pronoun, age group, education, and previous experience in visual experiment were also collected.

@@

Two types of departures were considered with the corresponding data generating process being designed for Experiment I and II.

@@

```{=tex}
\begin{equation} \label{eq:scaling-function}
\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
\end{equation}
```
where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.

@@

```{=tex}
\begin{align} \label{eq:heter-model}
\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}, 1 + 2 - |a| b (\boldsymbol{x} - a)^2 \boldsymbol{I}), \\
\end{align}
```
where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vector of size $n$.

The null regression model used to fit the realizations generated by the above model is formulated exactly the same as the model used in the first experiment:

```{=tex}
\begin{equation}
\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
\end{equation}
```
where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.

@@

### Factors

-   Treatments for nonlinearity experiment (plots summarising levels)
-   Treatments for heteroskedasticity experiment (plots summarising levels)

Experiment data were simulated using four different order of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$), three different sample sizes (n = 50, 100, 300) , four different standard deviations of the error ($\sigma$ = 0.5, 1, 2, 4) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$. A summary of the parameters used in this experiment is given in Table \ref{tab:parameter-table}.

The values of $j$ was chosen so that different shapes of non-linearity were included in the residual plot. These include "U" shape, "S" shape, "M" shape and "Triple-U" shape.

```{r different-shape-of-herimite}
stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1

@@

The range of $\sigma$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.

```{r different-sigma}
ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)

@@

Four different distribution were used to generate $X_{raw}$. The uniform and the normal distribution are symmetric and commonly assumed in statistical models. The adjusted log-normal distribution provides skewed density. And the discrete uniform distribution provides discreteness in residual plot, which could enrich the pool of visual patterns.

```{r different-dist}
ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)

@@

```{r results='asis'}
data.frame(n = c(50, 100, 300, ""),
           j = c(2, 3, 6, 18),
           sigma = c(0.5, 1, 2, 4),
           x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$")) %>%
  knitr::kable(col.names = NULL,
               escape = FALSE,
               format = "latex",
               booktabs = TRUE, 
               caption  = 'Parameter values for $n$, $j$ $\\sigma$, $X_{raw}$', 
               label = "parameter-table" ) %>%
  kableExtra::add_header_above(c("Sample size\n(sym3)", "Order of Hermite polynomial\n(sym1)", "Error SD\n(sym2)", "Distribution of sym4\n")) %>%
  as.character() %>%
  gsub("sym1", "$j$", .) %>%
  gsub("sym2", "$\\\\sigma$", .) %>%
  gsub("sym3", "$n$", .) %>%
  gsub("sym4", "$X_{raw}$", .) %>%
  cat()
  # gsub("\\\\\\$", "$", .) %>%
  # cat()
```

```{r example-lineup, fig.cap="Example lineup \\label{fig:example-lineup}"}
VI_MODEL$plot_lineup(polynomials_lineup[[23]]$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(),
                     remove_axis = TRUE)
```

@@

>>>>

---
title: |
  Why aren't significance tests commonly used for linear regression diagnostics?
type: ARTICLE TEMPLATE
author:
  - name: Weihao Li
    affil: a
    email: weihao.li@monash.edu
  - name: Dianne Cook
    affil: a
    email: dicook@monash.edu
  - name: Emi Tanaka
    affil: a
    email: emi.tanaka@monash.edu
affiliation:
  - num: a
    address: |
      Department of Econometrics and Business Statistics, Monash University, Clayton, VIC, Australia
bibliography: paper.bib
abstract: |
  Abstract to fill.
keywords: |
  data visualization; visual inference; hypothesis testing; residual plots;
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
  \newcommand*{\Perm}[2]{{}^{#1}\!P_{#2}}%
output: rticles::tf_article
editor_options: 
  chunk_output_type: inline
---

@@

During the experiment, every subject was presented with a block of 20 lineups. For each lineup, the actual data plot was drawn as a standard residual plot of the null model with raw residuals on the y-axis and fitted values on the x-axis. An additional horizontal red line was added at $y = 0$ as a helping line. The 19 null datasets were generated by the residual rotation technique, and plotted in the same way. The lineup consisted of 20 residual plots with one randomly placed actual data plot. And for every lineup, the subject was asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots were from others. If there was no noticeable difference between plots in a lineup, subjects were permitted to select zero plots without providing the reason. No subject was shown the same lineup twice. Information about preferred pronoun, age group, education, and previous experience in visual experiment were also collected.

@@

Two types of departures were considered with the corresponding data generating process being designed for experiment I and II.

@@

```{=tex}
\begin{equation} \label{eq:null-model}
\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
\end{equation}
```
where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.

@@

```{=tex}
\begin{align} \label{eq:heter-model}
\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
\boldsymbol{x} = g(\boldsymbol{x}_{raw}, 1)\\
\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}, 1 + 2 - |a| b (\boldsymbol{x} - a)^2 \boldsymbol{I}), \\
\end{align}
```
where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vector of size $n$ and $g(.)$ is the scaling function defined in \ref{eq:scaling-function}.

The null regression model used to fit the realizations generated by the above model is formulated exactly the same as Equation \ref{eq:null-model}.

@@

### Treatments

> Experiment I

-   Treatments for nonlinearity experiment (plots summarising levels)

Data of experiment I were simulated using four different order of probabilist's Hermite polynomials ($j = 2, 3, 6, 18$), three different sample sizes (n = 50, 100, 300), four different standard deviations of the error ($\sigma$ = 0.5, 1, 2, 4) and four different distribution of $X_{raw}$: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$. A summary of the parameters used in this experiment is given in Table \ref{tab:parameter-table}.

```{r results='asis'}
data.frame(n = c(50, 100, 300, ""),
           j = c(2, 3, 6, 18),
           sigma = c(0.5, 1, 2, 4),
           x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$")) %>%
  knitr::kable(col.names = NULL,
               align = "rrrl",
               escape = FALSE,
               format = "latex",
               booktabs = TRUE, 
               caption  = 'Parameter values for $n$, $j$ $\\sigma$, $X_{raw}$', 
               label = "parameter-table" ) %>%
  kableExtra::add_header_above(c("Sample size\n(sym3)", "Order of Hermite polynomial\n(sym1)", "Error SD\n(sym2)", "Distribution of sym4\n")) %>%
  as.character() %>%
  gsub("sym1", "$j$", .) %>%
  gsub("sym2", "$\\\\sigma$", .) %>%
  gsub("sym3", "$n$", .) %>%
  gsub("sym4", "$X_{raw}$", .) %>%
  cat()
  # gsub("\\\\\\$", "$", .) %>%
  # cat()
```

The values of $j$ was chosen so that different shapes of non-linearity were included in the residual plot. As shown in Figure \ref{fig:different-shape-of-herimite}, these include "U" shape, "S" shape, "M" shape and "Triple-U" shape.

```{r different-shape-of-herimite, fig.cap="Example residual plots of the linear regression model used in experiment I. Four different values of $j$ are included in the experiment results in four different shapes of residuals."}
stand_dist <- function(x) (x - min(x))/max(x - min(x)) * 2 - 1

@@

The range of $\sigma$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous. The example residual plots are given in Figure \ref{fig:different-sigma}.

```{r different-sigma, fig.cap="Example residual plots of the linear regression model used in experiment I. Four different values of $\\sigma$ are included in the experiment to control the strength of the signal."}
ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)

@@

Four different distribution were used to generate $X_{raw}$ shown in Figure \ref{fig:different-dist}. The uniform and the normal distribution are symmetric and commonly assumed in statistical models. The adjusted log-normal distribution provides skewed density. And the discrete uniform distribution provides discreteness in residual plot, which could enrich the pool of visual patterns.

```{r different-dist, fig.cap="Example residual plots of the linear regression model used in experiment I. Four different distribution of $x_{raw}$ are used in the experiment to provide various visual patterns."}
ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)

@@

Figure \ref{fig:example-lineup} shows one of the lineups used in experiment I. This lineup was produced under $n = 300$, $j = 6$, $\sigma = 1$ and $X_{raw} \sim N(0.0.3^2)$. The actual data plot location was four. All five subjects correctly identified the actual data plot for this lineup.

```{r example-lineup, fig.cap="Lineup 24 in experiment I. Can you spot the most different plot? \\label{fig:example-lineup}"}
VI_MODEL$plot_lineup(polynomials_lineup[[24]]$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(),
                     remove_axis = TRUE)
```

@@

> Experiment II

-   Treatments for heteroskedasticity experiment (plots summarising levels)

### 2022-09-21 paper_comparison/paper_comparison/paper_comparison.tex
update exp design for exp I

During the experiment, every subject was presented with a block of 20
lineups. And for every lineup, the subject was asked to select one or
more plots that are most different from others, provide a reason for
their selections, and evaluate how different they think the selected
plots were from others. If there was no noticeable difference between
plots in a lineup, subjects were permitted to select zero plots without
providing the reason. No subject was shown the same lineup twice.
Information about preferred pronoun, age group, education, and previous
experience in visual experiment were also collected.

@@

\begin{equation} \label{eq:scaling-function}
\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
\end{equation}

@@

\begin{align} \label{eq:heter-model}
\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}, 1 + 2 - |a| b (\boldsymbol{x} - a)^2 \boldsymbol{I}), \\
\end{align}

where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
\(\boldsymbol{\varepsilon}\) are vector of size \(n\).

The null regression model used to fit the realizations generated by the
above model is formulated exactly the same as the model used in the
first experiment:

\begin{equation}
\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
\end{equation}

@@

\hypertarget{factors}{%
\subsubsection{Factors}\label{factors}}

\begin{itemize}
\tightlist
\item
  Treatments for nonlinearity experiment (plots summarising levels)
\item
  Treatments for heteroskedasticity experiment (plots summarising
  levels)
\end{itemize}

Experiment data were simulated using four different order of
probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)), three different
sample sizes (n = 50, 100, 300) , four different standard deviations of
the error (\(\sigma\) = 0.5, 1, 2, 4) and four different distribution of
\(X_{raw}\): (1) \(U(-1, 1)\), (2) \(N(0, 0.3^2)\), (3)
\(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). A summary of the
parameters used in this experiment is given in Table
\ref{tab:parameter-table}.

The values of \(j\) was chosen so that different shapes of non-linearity
were included in the residual plot. These include ``U'' shape, ``S''
shape, ``M'' shape and ``Triple-U'' shape.

\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-shape-of-herimite-1}

The range of \(\sigma\), which is a factor controlling the strength of
the signal, was chosen so that different difficulty levels of lineups
were generated, and therefore, the estimated power curve would be smooth
and continuous.

\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-sigma-1}

Four different distribution were used to generate \(X_{raw}\). The
uniform and the normal distribution are symmetric and commonly assumed
in statistical models. The adjusted log-normal distribution provides
skewed density. And the discrete uniform distribution provides
discreteness in residual plot, which could enrich the pool of visual
patterns.

\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-dist-1}

@@

\caption{\label{tab:parameter-table}Parameter values for $n$, $j$ $\sigma$, $X_{raw}$}
\centering
\begin{tabular}[t]{lrrl}
\toprule
\multicolumn{1}{c}{\makecell[c]{Sample size\\($n$)}} & \multicolumn{1}{c}{\makecell[c]{Order of Hermite polynomial\\($j$)}} & \multicolumn{1}{c}{\makecell[c]{Error SD\\($\sigma$)}} & \multicolumn{1}{c}{\makecell[c]{Distribution of $X_{raw}$\\}} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4}
50 & 2 & 0.5 & $U(-1, 1)$\\
100 & 3 & 1.0 & $N(0, 0.3^2)$\\
300 & 6 & 2.0 & $lognormal(0, 0.6^2)/3$\\
 & 18 & 4.0 & $U\{1, 5\}$\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}
\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/example-lineup-1} \caption{Example lineup \label{fig:example-lineup}}\label{fig:example-lineup}
\end{figure}

>>>>

During the experiment, every subject was presented with a block of 20
lineups. For each lineup, the actual data plot was drawn as a standard
residual plot of the null model with raw residuals on the y-axis and
fitted values on the x-axis. An additional horizontal red line was added
at \(y = 0\) as a helping line. The 19 null datasets were generated by
the residual rotation technique, and plotted in the same way. The lineup
consisted of 20 residual plots with one randomly placed actual data
plot. And for every lineup, the subject was asked to select one or more
plots that are most different from others, provide a reason for their
selections, and evaluate how different they think the selected plots
were from others. If there was no noticeable difference between plots in
a lineup, subjects were permitted to select zero plots without providing
the reason. No subject was shown the same lineup twice. Information
about preferred pronoun, age group, education, and previous experience
in visual experiment were also collected.

@@

\begin{equation} \label{eq:null-model}
\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
\end{equation}

@@

\begin{align} \label{eq:heter-model}
\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
\boldsymbol{x} = g(\boldsymbol{x}_{raw}, 1)\\
\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}, 1 + 2 - |a| b (\boldsymbol{x} - a)^2 \boldsymbol{I}), \\
\end{align}

where \(\boldsymbol{y}\), \(\boldsymbol{x}\),
\(\boldsymbol{\varepsilon}\) are vector of size \(n\) and \(g(.)\) is
the scaling function defined in \ref{eq:scaling-function}.

The null regression model used to fit the realizations generated by the
above model is formulated exactly the same as Equation
\ref{eq:null-model}.

@@

\hypertarget{treatments}{%
\subsubsection{Treatments}\label{treatments}}

\begin{quote}
Experiment I
\end{quote}

\begin{itemize}
\tightlist
\item
  Treatments for nonlinearity experiment (plots summarising levels)
\end{itemize}

Data of experiment I were simulated using four different order of
probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)), three different
sample sizes (n = 50, 100, 300), four different standard deviations of
the error (\(\sigma\) = 0.5, 1, 2, 4) and four different distribution of
\(X_{raw}\): (1) \(U(-1, 1)\), (2) \(N(0, 0.3^2)\), (3)
\(lognormal(0, 0.6^2)/3\) and (4) \(u\{1, 5\}\). A summary of the
parameters used in this experiment is given in Table
\ref{tab:parameter-table}.

@@

\caption{\label{tab:parameter-table}Parameter values for $n$, $j$ $\sigma$, $X_{raw}$}
\centering
\begin{tabular}[t]{rrrl}
\toprule
\multicolumn{1}{c}{\makecell[c]{Sample size\\($n$)}} & \multicolumn{1}{c}{\makecell[c]{Order of Hermite polynomial\\($j$)}} & \multicolumn{1}{c}{\makecell[c]{Error SD\\($\sigma$)}} & \multicolumn{1}{c}{\makecell[c]{Distribution of $X_{raw}$\\}} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4}
50 & 2 & 0.5 & $U(-1, 1)$\\
100 & 3 & 1.0 & $N(0, 0.3^2)$\\
300 & 6 & 2.0 & $lognormal(0, 0.6^2)/3$\\
 & 18 & 4.0 & $U\{1, 5\}$\\
\bottomrule
\end{tabular}
\end{table}

The values of \(j\) was chosen so that different shapes of non-linearity
were included in the residual plot. As shown in Figure
\ref{fig:different-shape-of-herimite}, these include ``U'' shape, ``S''
shape, ``M'' shape and ``Triple-U'' shape.

\begin{figure}
\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-shape-of-herimite-1} \caption{Example residual plots of the linear regression model used in experiment I. Four different values of $j$ are included in the experiment results in four different shapes of residuals.}\label{fig:different-shape-of-herimite}
\end{figure}

The range of \(\sigma\), which is a factor controlling the strength of
the signal, was chosen so that different difficulty levels of lineups
were generated, and therefore, the estimated power curve would be smooth
and continuous. The example residual plots are given in Figure
\ref{fig:different-sigma}.

\begin{figure}
\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-sigma-1} \caption{Example residual plots of the linear regression model used in experiment I. Four different values of $\sigma$ are included in the experiment to control the strength of the signal.}\label{fig:different-sigma}
\end{figure}

Four different distribution were used to generate \(X_{raw}\) shown in
Figure \ref{fig:different-dist}. The uniform and the normal distribution
are symmetric and commonly assumed in statistical models. The adjusted
log-normal distribution provides skewed density. And the discrete
uniform distribution provides discreteness in residual plot, which could
enrich the pool of visual patterns.

\begin{figure}
\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-dist-1} \caption{Example residual plots of the linear regression model used in experiment I. Four different distribution of $x_{raw}$ are used in the experiment to provide various visual patterns.}\label{fig:different-dist}
\end{figure}

Figure \ref{fig:example-lineup} shows one of the lineups used in
experiment I. This lineup was produced under \(n = 300\), \(j = 6\),
\(\sigma = 1\) and \(X_{raw} \sim N(0.0.3^2)\). The actual data plot
location was four. All five subjects correctly identified the actual
data plot for this lineup.

\begin{figure}
\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/example-lineup-1} \caption{Lineup 24 in experiment I. Can you spot the most different plot? \label{fig:example-lineup}}\label{fig:example-lineup}
\end{figure}

\begin{quote}
Experiment II
\end{quote}

\begin{itemize}
\tightlist
\item
  Treatments for heteroskedasticity experiment (plots summarising
  levels)
\end{itemize}
